Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-07-25 doc/PRF2022/PRF_main.tex
yny

\begin{abstract}
Active colloidal systems with non-equilibrium self-organization is a long-standing challenge in biology. To understand how hydrodynamic flow may be used
to actively control self-assembly of Janus particles (JPs),  we use a model recently developed for the many-body hydrodynamics of amphiphilic JPs suspended 
in a viscous background flow (JFM, 941, 2022). We investigate how various bilayer structures arise from tuning the hydrophobicity/hydrophilicity of the JPs. 
Focusing on three distributions of hydrophobicity we found JPs may assemble into uni-lamella, multi-lamella, and striated structures in a viscous fluid. 
Under a linear flow and a Taylor-Green mixing flow,  we use three measures to quantify the collective dynamics of JP particles under a background flow: 
(a) Free energy from the hydrophobic interactions between the JPs, (b) an order parameter for the ordering of JPs in terms of alignment of their directors, 
and (c) a strain parameter that captures the deformation in the assembly. We found the dynamics of these three measures correlate well with the hydrodynamics 
of JPs. These numerical provide insights into dynamic control of non-equilibrium active biological systems with similar self-organization.
% insert abstract here
\end{abstract}

@@

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{align}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.
Without inertia, the system in~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
also satisfies the force-free and torque-free conditions.
%This means
%that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
%stress on each particle (with subscript `$i$') balance the force and
%torque from the hydrophobic potential~\cite{Fu2022_JFM}. 
The total force
and torque from the fluid flow are
\begin{align}
  \label{eq:force}
 \FF_i = \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i = \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{align}
respectively, where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu +
\nabla \uu^T \right)$ is the hydrodynamic stress tensor (pressure
tensor), and $\nnu_i$ is the unit outward normal on the $i$th particle.

The total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
stress on each particle (with subscript `$i$') balance the force and
torque from the hydrophobic potential. 
%$\FF_i$ and $T_i$ are to be in balance with the force and torque from the hydrophobic interactions between the
%particles and the viscous solvent. 
As in~\citet{Fu2022_JFM} such a hydrophobic attraction potential
$u(\xx,t)$ is assumed to the minimizer of the free energy
\begin{align}
\label{eq:free_energy}
F = \gamma
  \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} u^2 \right)
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{align}
where the hydrophobic attraction potential $u(\xx,t)$ can be regarded as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
$c$ is an ``effective" distance between particle centers, below which the steric repulsion between a pair of JPs becomes important.
In the followings we let $c$ be the radius of each JP.

The governing equation for $u(\xx,t)$ is obtained by computing the
domain variation of the free energy $F$~\cite{Fu20}. We obtain a
screened Laplace equation boundary value problem as a model for the
hydrophobic interaction between water and JPs:
\begin{alignat}{3}
  \label{eq:SLpde}
  -\rho^2 \Delta u + u &= 0,\qquad &&\xx \in \Omega, \\
  \label{eq:SLbc}
  u &= g, &&\xx \in \bd\Omega, \\
  \label{eq:SLfarfield}
  u &\rightarrow 0, &&|\xx| \rightarrow \infty,
\end{alignat}
where the boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

The hydrophobic stress is
\begin{align}
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
\end{align}
and is derived by applying variation of domain techniques to the free
energy~\eqref{eq:free_energy}~\cite{Fu20}. Then, the hydrophobic force
and torque are
\begin{align}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{align}
which balance the hydrodynamic force and torque~\eqref{eq:force}, for $i
= 1,\dots,N_b$. 
%The term 
%\begin{align}
%\mathbf{T}
%= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
%  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
%\end{align}
%is the hydrophobic stress~\cite{Fu20} and derived by variation of the
%domain techniques.

Finally, the dimensionless repulsion profile $P$ takes the form $P(s) =
1 - \sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$.  As such, only those particles within a distance $\rho_0$ of
$\Gamma_i$ contribute to repulsion. The parameter $M$ is the repulsion
modulus.

We use the same parameters from previous work~\cite{Fu2022_JFM}: $\mu =
1$ cP, $\gamma = 4$ pN nm$^{-1}$, $\rho = 5$ nm, $c = 1.25$ nm, $M = 2$
pN, and $\rho_0 = 0.5$ nm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral represenation and time stepping}
We recast~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
and~\eqref{eq:SLpde}--\eqref{eq:SLfarfield} as boundary integral
equations (BIEs) and discretize each BIE at $N$ points on each of the
$N_b$ particles with a collocation method. To express the solution
of~\eqref{eq:SLpde}--\eqref{eq:SLfarfield}, we adopt the double layer
potential
\begin{align}
\label{eq:SL_BIE}
  u({\xx}) = \DD[\sigma](\xx) = \int_\Gamma 
    \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, 
    \dif s_\yy, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental
solution to the screened Laplace equation~\eqref{eq:SLpde}, $K_0$ is the
zeroth-order modified Bessel function of the first kind, and $\sigma$ is
a scalar-valued density function.  The subscript in $\dif s_\yy$ denotes
integration with respect to $\yy \in \Gamma$.  For the velocity, we use
the representation~\cite{pow-mir1987}
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$, Stokeslet
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr
\otimes \rr}{|\rr|^2}\right)$, and Rotlet $\RR(\xx,\aa_i) =
\frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$, $\rr = \xx - \aa_i$.

@@

First, we must simplify the form of the free energy
\eqref{eq:free_energy} using integration by parts and \eqref{eq:SLpde}
to obtain
\begin{equation}
\label{eq:free_energy2}
F = -\gamma
\int_{\Gamma} \rho g \nabla u \cdot \nnu \,\dif s
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right).
\end{equation}
%
Here, we have substituted $g$ for $u$ since the boundary values are given.
However, evaluating $\nabla u \cdot \nnu$ on $\Gamma$ based on \eqref{eq:SL_BIE}
involves calculating a quadruple layer potential which has a
well-known obstacle in numerical implementation.
To overcome this obstacle, we use
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu(\xx)=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \Gamma.
\end{equation}
%
Here, ${\bf t}(\xx)$ is the tangent vector and $\dif/\dif s$ is the arclength derivative.
See Appendix in \S\ref{sec:appendix} for the proof and details.
Substituting \eqref{eq:normal_deriv} into \eqref{eq:free_energy2} for the normal derivative
leads to single layer integral which is more straightforward to evaluate and we compute the
arclength derivative using spectrally accurate Fourier transform.
We have validated the form of \eqref{eq:normal_deriv} by extrapolating 
the primal energy \eqref{eq:free_energy} along test curves placed slightly
outside the particles.

>>>>

\begin{abstract}
Active colloidal systems with non-equilibrium self-organization is a long-standing challenge in biology. To understand how hydrodynamic flow may be used
to actively control self-assembly of Janus particles (JPs),  we use a model recently developed for the many-body hydrodynamics of amphiphilic JPs suspended 
in a viscous background flow (JFM, 941, 2022) to investigate how various bilayer structures arise from tuning the hydrophobicity/hydrophilicity of the JPs. 
Focusing on three distributions of hydrophobicity we found JPs may assemble into uni-lamella, multi-lamella, and striated structures in a viscous fluid. 
Under a linear flow and a Taylor-Green mixing flow,  we use three measures to quantify the collective dynamics of JP particles under a background flow: 
(a) Free energy from the hydrophobic interactions between the JPs, (b) an order parameter for the ordering of JPs in terms of alignment of their directors, 
and (c) a strain parameter that captures the deformation in the assembly. We found the dynamics of these three measures correlate well with the hydrodynamics 
of JPs. These numerical provide insights into dynamic control of non-equilibrium active biological systems with similar self-organization.
% insert abstract here
\end{abstract}

@@

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time. The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis. The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{equation}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{equation}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$. 
Without inertia, the system in \eqref{eq:stokes}-\eqref{eq:rigid-vel} also satisfies the force-free and torque-free conditions.
This means that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic stress on each particle (with subscript `$i$')
are in balance with the force and torque from the hydrophobic potential \cite{Fu2022_JFM}.
The total force and torque from the fluid flow are computed as
\begin{equation}
  \label{eq:force}
 \FF_i \equiv \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i \equiv \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{equation}
where 
%$\FF_i \in \mathbb{R}^2$ and $T_i \in \mathbb{R}$
%are interparticle forces and torques. 
$\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor), and
$\nnu_i$ is the unit outward normal on the $i$th particle.

$\FF_i$ and $T_i$ are to be in balance with the force and torque from the hydrophobic interactions between the
particles and the viscous solvent. As in \citet{Fu2022_JFM} such a hydrophobic attraction potential $u(\xx,t)$ is assumed to
the minimizer of the free energy
\begin{equation}
\label{eq:free_energy}
F = \gamma
\int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{equation}
where the hydrophobic attraction potential $u(\xx,t)$ can be regarded as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
$c$ is an ``effective" distance between particle centers, below which the steric repulsion between a pair of JPs becomes important.
In the followings we use the radius of JP for $c$.

The governing equation for $u(\xx,t)$ is obtained by computing the domain variation of the free energy $F$ \cite{Fu20}:
We obtain
a screened Laplace equation boundary value problem as a model for the hydrophobic interaction between water and JPs:
\begin{gather}
  \label{eq:SL}
  -\rho^2 \Delta u + u = 0,\quad \xx \in \Omega, \\
  u = g, \quad \xx \in \bd\Omega, \quad 
  u \rightarrow 0, \quad |\xx| \rightarrow \infty.
\end{gather}
With 
The boundary condition $g$ gives the degree of hydrophobicity the
particle-solvent interface.

Variation of the the free energy \eqref{eq:free_energy} provides
\begin{equation}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{equation}
for $i = 1,\dots,N_b$. The term 
$
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
$
is the hydrophobic stress identified in ~\cite{Fu20} and derived by
variation of the domain techniques.

The dimensionless repulsion profile $P$ takes the form $P(s) = 1 -
\sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$.  As such, only those particles within a distance $\rho_0$ of
$\Gamma_i$ contribute to repulsion.  The parameter $M$ is the repulsion
modulus.

We use the parameters $\mu = 1$ cP, $\gamma = 4$ pN nm$^{-1}$,
$\rho = 5$ nm, $c = 1.25$ nm, $M = 2$ pN, and $\rho_0 = 0.5$ nm.

\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}--\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL},
$K_0$ is the zeroth-order modified Bessel function of the first kind,
and $\sigma$ is a scalar-valued density function.
The subscript in $\dif s_\yy$ denotes integration with respect to $\yy \in \Gamma$. 
For the velocity, we use the representation [REF]
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$,
$\rr = \xx - \aa_i$ [REF].

@@

First, we must simplify the
form of the free energy \eqref{eq:free_energy}
using integration by parts and \eqref{eq:SL}
to obtain
\begin{equation}
\label{eq:free_energy2}
F = -\gamma
\int_{\Gamma} \rho g \nabla u \cdot \nnu \,\dif s
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right).
\end{equation}
%
Here, we have substituted $g$ for $u$ since the boundary values are given.
However, evaluating $\nabla u \cdot \nnu$ on $\Gamma$ based on \eqref{eq:SL_BIE}
involves calculating a quadruple layer potential which has a
well-known obstacle in numerical implementation.
To overcome this obstacle, we use
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu(\xx)=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \Gamma.
\end{equation}
%
Here, ${\bf t}(\xx)$ is the tangent vector and $\dif/\dif s$ is the arclength derivative.
See Appendix in \S\ref{sec:appendix} for the proof and details.
Substituting \eqref{eq:normal_deriv} into \eqref{eq:free_energy2} for the normal derivative
leads to single layer integral which is more straightforward to evaluate and we compute the
arclength derivative using spectrally accurate Fourier transform.
We have validated the form of \eqref{eq:normal_deriv} by extrapolating 
the primal energy \eqref{eq:free_energy} along test curves placed slightly
outside the particles.
